Abstract

### 2025-02-14 output/Letter_to_editor_reviewers.Rmd
Version last response to reviewer 2

@@

@@

> I was at a loss at the end of the discussion to identify the interpretive or methodological value that was gained by working with non-orthogonal axes.
> The correlation between items can be represented by the angle between item vectors in the space with orthogonal axes, and the correlation between coordinates for persons can be represented by the shape of the scatter plot of points representing the individuals.
> How are these features interpreted when the coordinate axes are non-orthogonal?

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

The reviewer is right in this respect; we made the mistake of assuming this relationship was obvious enough, but we acknowledge it may be confusing to the reader to refer to the $MDISC_i$ and obviate the influence of $d_i$. 
We have added a clarification in this respect to the end of the paragraph, which now reads:

    > That is, the less discriminating the item is,
    the further away it will be from the origin.
    Of course, it is also directly proportional to
    the intercept parameter $`r INTERCEPT_PARAM`$.
    This implies that, if $`r INTERCEPT_NULL_EQ`$,
    $`r DISTANCE_NULL_EQ`$ as well,
    regardless of the discrimination.

@@

Although we argue indeed that the coordinate system is arbitrary, our stance is also that there is a more interpretable approach than simply assuming orthogonal axes.
This approach implies that there is a geometric basis (and thus a coordinate system) that can be considered a faithful representation of the covariance structure.

@@

We are glad that the reviewer has found our work thought-provoking.
We argue indeed that different samples may yield different latent covariance structures.
This problem has received much attention in the factor analysis literature (i.e., factorial invariance), it has been largely overlooked in the MIRT literature, where any covariance structure is usually represented in orthogonal axes (implying an orthonormal basis).

> [...] I was at a loss at the end of the discussion to identify the interpretive or methodological value that was gained by working with non-orthogonal axes.
> The correlation between items can be represented by the angle between item vectors in the space with orthogonal axes, and the correlation between coordinates for persons can be represented by the shape of the scatter plot of points representing the individuals.
> How are these features interpreted when the coordinate axes are non-orthogonal?

We believe that non-orthogonal axes provide a faithful representation of the latent covariance structure:
In geometrical terms, the cosine between axes directly maps to the correlation between the two dimensions represented by those axes.
In the orthogonal space, the angles between items are "distorted" and do not faithfully represent the correlation among those items.
Consider the following example: Two simple items, each measuring a single dimension of a two-dimensional construct, will always be represented by a 90-degree angle in the orthogonal coordinate system.
If those two dimensions are uncorrelated, so the items will also be.
However, suppose those two dimensions are highly correlated; in that case, the items would also have to be equally highly correlated, but would still be represented by a 90-degree angle in orthogonal coordinates.
Therefore, our argument is that the correlation among items is better represented by the cosines of their vectors in the non-orthogonal space.

Regarding the person parameters, the interpretation is more complicated:
If the original, correlated coordinates of the persons are represented in the orthogonal space, the shape of the scatter plot indeed represents the correlation between the latent trait coordinates.
However, our rationale is that a basis may be used such that, when the coordinates are transformed to an orthonormal one, the resulting scatter plot is spherical (i.e. its covariance is an identity matrix).
This results in a basis (and an associated coordinate system) that does not have a direct interpretation, save for being the "inverse transpose" of the basis where we argue the items are faithfully represented.

@@

The distances between points are given by the norm, which we argue to be invariant to changes of bases.
Interestingly, when considering the covariance-based version of the multidimensional parameters, the distance between points in the latent space is given by the Mahalanobis distance.

@@

We apologize for the lack of clarity.
We intended to refer to footnote 1 in page 16, we have correctly referenced the page and footnote in the new version.

@@

We intended to refer to <!-- # TODO: Complete -->
Please note that, as we do not have access to the 2nd edition anymore, we have used the 3rd one in our new version, and updated all the references accordingly.

@@

This is the expression of a vector represented in a basis, in this case, representing that the vector can be represented in each of the two bases indistinctly.
Following the reviewer's suggestion, we have added this clarification to the text, so it now reads:

    > [...] such that $`r TRAIT_VECTOR_DEF_EQ`$.
    That is, vector $`r TRAIT_VECTOR`$ is represented in each of the two bases by
    the linear combination of its element vectors
    scaled by the coordinates of $`r TRAIT_VECTOR`$ in that basis.

@@

The -1 superscripts indeed indicate the inverse of a matrix.
This has been an unfortunate typesetting problem of the submission system,
that renders the -1 and the *T* at the same height.
We have sorted this out by using parentheses, as you will be able to see in our new submission.

@@

<!-- # TODO: Add response -->

@@

What we want to imply is that there are two different spaces, the latent space, and the test space, and that, unlike the latent traits, which "naturally lie" in the latent space, the item parameters "naturally lie in the test space", even though they can be projected onto the latent space.
This happens even when the basis of the latent space (and, consequently, that of the test space) is orthonormal, whih we argue is overlooked (due to a lack of need to make this distinction) in Reckase's original formulations.

@@

As we mentioned above, when the attributes are correlated, a distance measure that keeps the norm invariant in the latent space is the Mahalanobis distance.
One should make the distinction between representing the attributes being measured (i.e., the latent traits) and the variables measuring those attributes (in our case, the items).
It is when we represent the items that we should use axes that "faithfully represent the covariance structure", and not when we represent the attributes.
